---
title: "Homework_06"
author: "Shelby N. Scarfo"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

### Class Sample Data
```{r}
#Load packages
library(ggplot2) #for graphics
library(MASS) #for maximum likelihood estimation
library(dplyr)

#Read in data vector
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

#Plot histogram of data
p1 <- ggplot(data=z, aes(x=myVar, y=after_stat(density))) +
  geom_histogram(color="grey60",fill="paleturquoise1",linewidth=0.2) 
print(p1)

#Add empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

#Get maximum likelihodd parameters for normal
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

#Plot normal probability density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
 
#Plot exponential probability density
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
 
#Plot uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
 
#Plot gamma probability density
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 
#Plot beta probability density
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=after_stat(density))) +
  geom_histogram(color="grey60",fill="paleturquoise1",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

### Alewife length-frequency data
#### Alewife length-frequency data from trawling conducted from 2015-2021
```{r}
#Bringing in trawling data dn filtering by alewife, removing NAs, and selecting for only age 1 fish.

fish_lengths <- read.csv("Ind_fish_sampled_detailed.csv", header = TRUE, sep = ",")

alw_lengths_1 <- fish_lengths %>%
  filter(spec_code == "ALW") %>%
  filter(!tl_mm == "NA") %>%
  filter(tl_mm <= 120)

str(alw_lengths_1)
summary(alw_lengths_1)
```

```{r}
#Plot histogram of data
p1 <- ggplot(data = alw_lengths_1, aes(x = tl_mm, y = after_stat(density))) + 
  geom_histogram(color = "grey60", fill = "darkseagreen1", size = 0.2)
print(p1)
```

```{r}
#Add empirical density curve
p1 <- p1 + geom_density(linetype = "dotted", size = 0.75)
print(p1)
```

```{r}
#Get maximum likelihood parameters for normal
normPars <- fitdistr(alw_lengths_1$tl_mm, "normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```

```{r}
#Plot normal probability density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0, max(alw_lengths_1$tl_mm), len = length(alw_lengths_1$tl_mm))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour = "red", n = length(alw_lengths_1$tl_mm), args = list(mean = meanML, sd = sdML))
p1 + stat
```

```{r}
#Plot exponential probability density
expoPars <- fitdistr(alw_lengths_1$tl_mm, "exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour = "blue", n = length(alw_lengths_1$tl_mm),args = list(rate = rateML))

p1 + stat + stat2
```

```{r}
#Plot uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour = "darkgreen", n = length(alw_lengths_1$tl_mm), args = list(min= min(alw_lengths_1$tl_mm), max= max(alw_lengths_1$tl_mm)))

p1 + stat + stat2 + stat3
```

```{r}
#Plot gamma probabiliuty density
gammaPars <- fitdistr(alw_lengths_1$tl_mm, "gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour = "brown", n = length(alw_lengths_1$tl_mm), args = list(shape = shapeML, rate = rateML))

p1+ stat +stat2 + stat3 + stat4
```

```{r}
#Plot beta probability density
pSpecial <- ggplot(data = alw_lengths_1, aes(x = tl_mm/(max(tl_mm +0.1)), y = after_stat(density)))+
  geom_histogram(color = "grey60", fill = "darkseagreen1", size = 0.2)+
  xlim(c(0,1))+
  geom_density(size = 0.75, linetype = "dotted")
betaPars <- fitdistr(x = alw_lengths_1$tl_mm/max(alw_lengths_1$tl_mm +0.1), start = list(shape1 = 1, shape2 = 2), "beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour = "orchid",n = length(alw_lengths_1$tl_mm), args = list(shape1 = shape1ML, shape2 = shape2ML))

pSpecial + statSpecial
```

### Explanation
A normal distribution fits this data set well, it is a fairly large dataset and it consists of the total length in millimeters of alewife <= age 1 from Lake Champlain. The age classes are usually represented in distinct total length bins within a species. It makes sense that the density of the mean value of the dataset would be the highest because that is that average length of an age-1 fish and the other lengths would be less common. 


### Simulated Data

```{r}
z <- rnorm(n = 1961, mean = 74.6746558, sd = 16.5137192)
z <- data.frame(1:1961, z)
names(z) <- list("ID", "myVar")
z <- z[z$myVar >0,]
str(z)
summary(z$myVar)


normPars <- fitdistr(z$myVar, "normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]


meanSim <- normPars$estimate["mean"]
sdSim <- normPars$estimate["sd"]

simVal <- seq(0, max(z$myVar), len = length(z$myVar))

p2 <- ggplot(data = z, aes(x = myVar, y = after_stat(density)))+
  geom_histogram(color = "grey60", fill = "thistle2", size = 0.2)
print(p2)

p2 <-  p2 +  geom_density(linetype="dotted",size=0.75)
print(p2)

stat_sim <- stat_function(aes(x = simVal, y = ..y..), fun = dnorm, colour = "red", n = length(z$myVar), args = list(mean = meanSim, sd = sdSim))
p2 + stat_sim
```

```{r}
#Final figures
print(p1+stat)
print(p2+stat_sim)
```


The original dataset is large enough that it is representative of the population. Using the same n, the actual and simulated distributions are very similar.
