---
title: "Homework_07"
author: "Shelby N. Scarfo"
date: "2024-02-28"
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am using a bottom trawling dataset that looks at the catch per unit effort (CPUE) of alewife throughout the three different portions (north, central, south) of the main lake in Lake Champlain. I am not expecting to see any significant differences in CPUE between these regions because they are all part of the main lake and are not geographically isolated from eachother.

### Running ANOVA and Tukey tests on my own data
```{r, message=FALSE, warning=FALSE}
#Libraries
library(ggplot2)
library(MASS)
library(dplyr)
library(truncnorm)
```

```{r}
#Loading in data and getting attributes

ALW_CPUE_basin <- read.csv("ALW_CPUE_basin.csv")

ALW_CPUE_basin$minor_basin <- factor(ALW_CPUE_basin$minor_basin, levels = c("North Main Lake", "Central Main Lake", "South Main Lake"))

str(ALW_CPUE_basin)

summary(ALW_CPUE_basin)

ALW_CPUE <- group_by(ALW_CPUE_basin, minor_basin) %>%
  summarise(
    count = n(),
    mean = mean(CPUE, na.rm = TRUE),
    sd = sd(CPUE, na.rm = TRUE)
  )
print(ALW_CPUE) # This shows the sample sizes, means, and variances for each group

```

```{r}
#Running ANOVA on data and creating figure

res.aov <- aov(CPUE~minor_basin, data = ALW_CPUE_basin)
summary(res.aov)

TukeyHSD(res.aov)

ANOplot <- ggplot(data = ALW_CPUE_basin)+
  aes(x= minor_basin, y = CPUE, fill = minor_basin) +
  geom_boxplot()
ANOplot

```

The ANOVA and tukey tests shows that there is no statistical difference (p > 0.05) between CPUE in the three regions of the main lake.

### Making fake dataset that has the same attributes as my real one

```{r}
# Giving this data the same parameters as my original data
nGroup <- 3
nName <- c("North Main Lake", "Central Main Lake", "South Main Lake")
nSize <- c(72, 218, 55)
nMean <- c(100.4, 90.1, 54.9)
nSD <- c(199.7, 169.8, 119.2)
ID <- 1:(sum(nSize))

#Used rtruncnorm to set my bottom limit to 0 because I can't have a negative catch
modelCPUE <- c(rtruncnorm(n=nSize[1], a = 0, b = Inf, mean=nMean[1],sd=nSD[1]),
            rtruncnorm(n=nSize[2], a = 0, b = Inf, mean=nMean[2],sd=nSD[2]),
            rtruncnorm(n=nSize[3], a = 0, b = Inf, mean=nMean[3],sd=nSD[3]))
TGroup <- rep(nName,nSize)
ANOdata <- data.frame(ID,TGroup,modelCPUE)

ANOdata$TGroup <- factor(ANOdata$TGroup, levels = c("North Main Lake", "Central Main Lake", "South Main Lake"))

str(ANOdata)
```


```{r}
#ANOVA and plot of fake data

ANOmodel <- aov(modelCPUE~TGroup,data=ANOdata)
summary(ANOmodel)

TestANOPlot <- ggplot(data=ANOdata,aes(x=TGroup,y=modelCPUE,fill=TGroup)) +
  geom_boxplot()
print(TestANOPlot)

TukeyHSD(ANOmodel) #This shows that there is no statistic difference between Central Main Lake CPUE and North Main Lake CPUE but there is a statistical difference between the South Main Lake CPUE and the other two regions.
```

### Altering the metrics of the fake data to see how the parameters affect statistical significance 

```{r}
#Double the mean of the South Main Lake fake data
set.seed(100)
nGroup <- 3
nName <- c("North Main Lake", "Central Main Lake", "South Main Lake")
nSize <- c(72, 218, 55)
nMean2 <- c(100.4, 90.1, 109.8)
nSD <- c(199.7, 169.8, 119.2)
ID <- 1:(sum(nSize))

meanDubCPUE <- c(rtruncnorm(n=nSize[1], a = 0, b = Inf, mean=nMean2[1],sd=nSD[1]),
            rtruncnorm(n=nSize[2], a = 0, b = Inf, mean=nMean2[2],sd=nSD[2]),
            rtruncnorm(n=nSize[3], a = 0, b = Inf, mean=nMean2[3],sd=nSD[3]))
TGroup <- rep(nName,nSize)
ANOMeanDubdata <- data.frame(ID,TGroup,meanDubCPUE)

ANOMeanDubdata$TGroup <- factor(ANOMeanDubdata$TGroup, levels = c("North Main Lake", "Central Main Lake", "South Main Lake"))

#Run ANOVA and Tukey on doubled South Main Lake mean data

ANOMeanDubmodel <- aov(meanDubCPUE~TGroup,data=ANOMeanDubdata)
summary(ANOMeanDubmodel)

ANOMeanDubPlot <- ggplot(data=ANOMeanDubdata,aes(x=TGroup,y=meanDubCPUE,fill=TGroup)) +
  geom_boxplot()
print(ANOMeanDubPlot)

TukeyHSD(ANOMeanDubmodel)


```
In order for there to be no statistical difference between the North Main Lake and the South Main Lake CPUEs I had to double the original mean of the South Main Lake.

```{r}
#Change the sample sizes
set.seed(110)
nGroup <- 3
nName <- c("North Main Lake", "Central Main Lake", "South Main Lake")
nSize2 <- c(72, 218, 15)
nMean <- c(100.4, 90.1, 54.9)
nSD <- c(199.7, 169.8, 119.2)
ID <- 1:(sum(nSize2))

sizeCPUE <- c(rtruncnorm(n=nSize2[1], a = 0, b = Inf, mean=nMean[1],sd=nSD[1]),
            rtruncnorm(n=nSize2[2], a = 0, b = Inf, mean=nMean[2],sd=nSD[2]),
            rtruncnorm(n=nSize2[3], a = 0, b = Inf, mean=nMean[3],sd=nSD[3]))
TGroup <- rep(nName,nSize2)
ANOSizedata <- data.frame(ID,TGroup,sizeCPUE)

ANOSizedata$TGroup <- factor(ANOSizedata$TGroup, levels = c("North Main Lake", "Central Main Lake", "South Main Lake"))

#Run ANOVA and Tukey on altered sample size data

ANOSizemodel <- aov(sizeCPUE~TGroup,data=ANOSizedata)
summary(ANOSizemodel)

ANOSizePlot <- ggplot(data=ANOSizedata,aes(x=TGroup,y=sizeCPUE,fill=TGroup)) +
  geom_boxplot()
print(ANOSizePlot)

TukeyHSD(ANOSizemodel)
```
In order for there to be no statistical difference between the Central Main Lake and the South Main Lake CPUEs (largest sample size difference) I had to reduce the South Main Lake sample size by 40 to 15.
